Parallel Hierarchical Clustering using \\ Rank-Two Nonnegative Matrix Factorization \thanks{Identify applicable funding agency here. If none, delete this.}

### 2020-07-15 latex/paper.tex
DC Mall Figure
included strong scaling plots for synthetic data
added description of communication model

\section{Preliminaries}
\label{sec:prelim}

\begin{itemize}
	\item NMF, BCD, NNLS
	\item PLANC
	\item communication model and collectives
\end{itemize}

@@

@@

\subsubsection{Hyperspectral Imaging}

>>>>

\section{Preliminaries and Related Work}
\label{sec:prelim}

\GB{use Kuang's related work for NMF (2016 paper for divide-and-conquer NMF in JGO: \url{https://link.springer.com/article/10.1007/s10898-017-0515-z}), use PPoPP/TKDE/arXiv of ours and also Liavas, pyramid NMF: \doi{10.1109/IJCNN.2003.1224004}}

\subsection{NMF}

\GB{general NMF, BCD/AO algorithms, NNLS}

\subsection{Parallel NMF}

\GB{PLANC and other related work}

\subsection{Communication Model}

We use the $\alpha$-$\beta$-$\gamma$ model \cite{TRG05,CH+07,BCDH+14} for analysis of distributed-memory parallel algorithms. 
In this model, the cost of sending a single message of $n$ words of data between two processors is $\alpha + \beta \cdot n$, so that $\alpha$ represents the latency cost of the message and $\beta$ represents the bandwidth cost of each word in the message.
The $\gamma$ parameter represents the computational cost of a single floating point operation (flop).
In this simplified communication model, we ignore contention in the network, assuming in effect a fully connected network, and other limiting factors in practice such as the number of hops between nodes and the network injection rate \cite{GOS16}.
We let $p$ represent the number of processors available on the machine.

All of the interprocessor communication in the algorithms presented in this work are encapsulated in collective communication operations that involve the full set of processors.
Algorithms for implementing the collective operations are built out of pairwise send and receive operations, and we assume the most efficient algorithms are used in our analysis \cite{TRG05,CH+07}.
The collectives used in our algorithms are all-reduce, all-gather, and reduce-scatter.
In an all-reduce, all processors start out with the same amount of data and all end with a copy of the same result, which is in our case a sum of all the inputs (and the same size as a single input).
The cost of an all-reduce of size $n$ words is $\alpha \cdot O(\log p) + (\beta+\gamma) \cdot O(n)$ for $n>p$ and $(\alpha+\beta) \cdot O(\log p)$ for $n<p$.
In an all-gather, all processors start out with separate data and all end with a copy of the same result, which is the union of all the input data.
If each processor starts with $n/p$ data and ends with $n$ data, the cost of the all-gather is $\alpha \cdot O(\log p) + \beta \cdot O(n)$.
In a reduce-scatter, all processors start out with the same amount of data and all end with a subset of the result, which is in our case a sum of all the inputs (and is smaller than its input).
If each processor starts with $n$ data and ends with $n/p$ data, the cost of the reduce-scatter is $\alpha \cdot O(\log p) + (\beta+\gamma) \cdot O(n)$.
In the case of all-reduce and reduce-scatter, the computational cost is typically dominated by the bandwidth cost because $\beta \gg \gamma$.

@@

\GB{track NNLS of $\M{W}$ term even though it's lower order}

@@

\subsubsection{Hyperspectral Imaging}
\begin{figure}
\input{fig/dc}
\caption{Hierarchical Cluster from DC Mall Image}
\label{fig:dc}
\end{figure}
